---
title: "Homework 3"
subtitle: 'FINM 33210: Bayesian Statistical Inference and Machine Learning'
author: "Ki Hyun"
date: 'Due: 23:59 (CT) May 19th 2023'
output: pdf_document
---

```{r packages, include=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
```


# Problem 0.1

## (a)

$$
\begin{aligned}
\mathbf{E}[\mathbf{h}' \mathbf{r}] &=
\mathbf{E}[\sum_{i = 1}^n h_i r_i] \\
&= \sum_{i = 1}^n \mathbf{E}[h_i r_i] \\
&= \frac{1}{n} \sum_{i = 1}^n \mathbf{E}[r_i] \\
&(\because \mathbf{h} = (1/n, 1/n, \cdots, 1/n)) \\
&= \frac{1}{n} \sum_{i = 1}^n \mathbf{E}[\beta r_M + \epsilon_i] \\
&= \frac{1}{n} \sum_{i = 1}^n \left( \mathbf{E}[\beta r_M] + 
\mathbf{E}[\epsilon_i] \right) \\
&= \frac{1}{n} \sum_{i = 1}^n \mathbf{E}[\beta r_M] +
\frac{1}{n} \sum_{i = 1}^n \mathbf{E}[\epsilon_i] \\
&= \frac{\beta}{n} \sum_{i = 1}^n \mathbf{E}[r_M] +
\frac{1}{n} \sum_{i = 1}^n \mathbf{E}[\epsilon_i] \\
&= \frac{\beta}{n} n \mathbf{E}[r_M]  +
\frac{1}{n} \sum_{i = 1}^n \mathbf{E}[\epsilon_i]\\
&= \beta \mathbf{E}[r_M] +
\frac{1}{n} \sum_{i = 1}^n \mathbf{E}[\epsilon_i]
\end{aligned}
$$

$$
\begin{aligned}
\mathbf{V}[\mathbf{h}' \mathbf{r}] &=
\mathbf{V}[\sum_{i = 1}^n h_i r_i] \\
&= \mathbf{V}[\frac{1}{n} \sum_{i = 1}^n r_i] \\
&(\because \mathbf{h} = (1/n, 1/n, \cdots, 1/n)) \\
&= \frac{1}{n^2} \mathbf{V}[\sum_{i = 1}^n r_i] \\
&= \frac{1}{n^2} \mathbf{V}[\sum_{i = 1}^n (\beta r_M + \epsilon_i)] \\
&= \frac{1}{n^2} \mathbf{V}[n \beta r_M + \sum_{i = 1}^n \epsilon_i] \\
&= \frac{1}{n^2} \left( \mathbf{V}[n \beta r_M] + 
\sum_{i = 1}^n \mathbf{V}[\epsilon_i] + 
\sum_{i = 1}^n 2Cov(n \beta r_M, \epsilon_i) +
\sum_{i \neq j} 2Cov(\epsilon_i, \epsilon_j)
\right) \\
&= \frac{1}{n^2} \left( (n \beta)^2 \mathbf{V}[r_M] + 
\sum_{i = 1}^n \mathbf{V}[\epsilon_i] + 
\sum_{i = 1}^n 2n \beta Cov(r_M, \epsilon_i) +
\sum_{i \neq j} 2Cov(\epsilon_i, \epsilon_j)
\right) \\
&= \frac{1}{n^2} \left( (n \beta)^2 \mathbf{V}[r_M] + 
\sum_{i = 1}^n \mathbf{V}[\epsilon_i] + 
2 \sum_{i \neq j} Cov(\epsilon_i, \epsilon_j)
\right) \\
&(\because \epsilon_i \perp \!\!\! \perp r_M) \\
&= \frac{1}{n^2} \left( (n \beta)^2 \mathbf{V}[r_M] + 
\sum_{i = 1}^n \mathbf{V}[\epsilon_i] \right) \\
&(\because \text{ for } i \neq j, \ \epsilon_i \perp \!\!\! \perp \epsilon_j) \\
&= \beta^2 \mathbf{V}[r_M] + 
\frac{1}{n^2} \sum_{i = 1}^n \mathbf{V}[\epsilon_i] \\
&= \beta^2 \sigma^2_M + \frac{1}{n^2} \sum_{i = 1}^n \sigma^2_i
\end{aligned}
$$

Therefore, the functions $f()$ and $g()$ can be explicitly defined as:

$$
f(\beta, \sigma^2_M) := \beta^2 \sigma^2_M
$$

and

$$
g(\sigma^2_1, \cdots, \sigma^2_n) := \frac{1}{n^2} \sum_{i = 1}^n \sigma^2_i
$$

\newpage

## (b)

It was given that $\beta = 0.5$, $\sigma_M = 0.2$, and $\sigma_i \approx 0.03$.

Using the definition in (a) for $f()$ and $g()$,

$$
\frac{f}{f + g} = \frac{\beta^2 \sigma^2_M}{\beta^2 \sigma^2_M + 
\frac{1}{n^2} \sum_{i = 1}^n \sigma^2_i}
$$

Now substituting the values  $\beta = 0.5$, $\sigma_M = 0.2$, and 
$\sigma_i \approx 0.03$.

$$
\frac{f}{f + g} (n) \approx \frac{0.5^2 \cdot 0.2^2}{
0.5^2 \cdot 0.2^2 + \frac{1}{n^2} n \cdot 0.03^2}
= \frac{0.01}{0.01 + \frac{0.0009}{n}}
= \frac{1}{1 + \frac{0.09}{n}}
$$

If we numerically compute and plot $f/(f + g)$ as a function of $n$ for 
$n = 2, \dots, 30$:

```{r q1b_plot, echo=FALSE}
f_over_fg <- function(n){
  1/(1 + 0.09/n)
}
n <- seq(2, 30, by = 1)

tibble(ffg = f_over_fg(n), n = n) %>% 
  ggplot(mapping = aes(x = n, y = ffg)) + 
  geom_point() + 
  geom_line() +
  ylim(c(0.95, 1)) +
  labs(x = "n", y = "f/(f + g)",
       title = "Fraction of Variance Explained by the Market") +
  theme_bw(base_size = 12)
```

\newpage

## (c)

We know from (a) that

$$
\mathbf{E}[\mathbf{h}' \mathbf{r}] = \beta \mathbf{E}[r_M] +
\frac{1}{n} \sum_{i = 1}^n \mathbf{E}[\epsilon_i]
$$

Therefore,

$$
\begin{aligned}
\mathbf{E}[\mathbf{h}' \mathbf{r} - 0.01] &=
\mathbf{E}[\mathbf{h}' \mathbf{r}] - \mathbf{E}[0.01] \\
&= \beta \mathbf{E}[r_M] + \frac{1}{n} \sum_{i = 1}^n \mathbf{E}[\epsilon_i]
- 0.01 \\
&= \beta \cdot 0.07 + \frac{1}{n} \sum_{i = 1}^n 1.5 \cdot \sigma_i - 0.01 \\
&= 0.07 \beta + \frac{1.5}{n} \sum_{i = 1}^n \sigma_i - 0.01
\end{aligned}
$$

Moreover, we also know from (a) that

$$
\mathbf{V}[\mathbf{h}' \mathbf{r}] = \beta^2 \sigma^2_M + 
\frac{1}{n^2} \sum_{i = 1}^n \sigma^2_i
$$

Now using the assumptions of (b)

$$
\begin{aligned}
\frac{\mathbf{E}[\mathbf{h}' \mathbf{r} - 0.01]}
{\sqrt{\mathbf{V}[\mathbf{h}' \mathbf{r}]}} &=
\frac{0.07 \beta + \frac{1.5}{n} \sum_{i = 1}^n \sigma_i - 0.01}
{\sqrt{\beta^2 \sigma^2_M + \frac{1}{n^2} \sum_{i = 1}^n \sigma^2_i}} \\
& \approx \frac{0.07 \cdot 0.5 + \frac{1.5}{n} \sum_{i = 1}^n 0.03 - 0.01}
{\sqrt{0.5^2 \cdot 0.2^2 + \frac{1}{n^2} \sum_{i = 1}^n 0.03^2}} \\
&= \frac{0.07 \cdot 0.5 + 1.5 \cdot 0.03 - 0.01}
{\sqrt{0.5^2 \cdot 0.2^2 + \frac{0.03^2}{n}}} \\
&= \frac{0.35 + 0.45 - 0.1}{\sqrt{1 + \frac{0.09}{n}}} \\
&= \frac{0.7}{\sqrt{1 + \frac{0.09}{n}}}
\end{aligned}
$$

If we numerically compute and plot the Sharpe ratio as a function of $n$ for 
$n = 2, \dots, 30$:

```{r q1c_plot_1, echo=FALSE}
sharpe1 <- function(n){
  0.7/sqrt(1 + 0.09/n)
}

tibble(sharpe.r = sharpe1(n), n = n) %>% 
  ggplot(mapping = aes(x = n, y = sharpe.r)) +
  geom_point() + 
  geom_line() +
  ylim(c(0.68, 0.7))+
  labs(x = "n", y = "Sharpe Ratio",
       title = "Sharpe Ratio of the fund-of-funds across n") +
  theme_bw(base_size = 12)
```

If the Sharpe ratio of $\epsilon_i$ is $2.0$,

$$
\begin{aligned}
\frac{\mathbf{E}[\mathbf{h}' \mathbf{r} - 0.01]}
{\sqrt{\mathbf{V}[\mathbf{h}' \mathbf{r}]}} &=
\frac{0.07 \beta + \frac{2.0}{n} \sum_{i = 1}^n \sigma_i - 0.01}
{\sqrt{\beta^2 \sigma^2_M + \frac{1}{n^2} \sum_{i = 1}^n \sigma^2_i}} \\
& \approx \frac{0.07 \cdot 0.5 + \frac{2.0}{n} \sum_{i = 1}^n 0.03 - 0.01}
{\sqrt{0.5^2 \cdot 0.2^2 + \frac{1}{n^2} \sum_{i = 1}^n 0.03^2}} \\
&= \frac{0.07 \cdot 0.5 + 2.0 \cdot 0.03 - 0.01}
{\sqrt{0.5^2 \cdot 0.2^2 + \frac{0.03^2}{n}}} \\
&= \frac{0.35 + 0.6 - 0.1}{\sqrt{1 + \frac{0.09}{n}}} \\
&= \frac{0.85}{\sqrt{1 + \frac{0.09}{n}}}
\end{aligned}
$$

\newpage

Moreover, if we plot the two different $\epsilon_i$-Sharpe-ratio together:

```{r q1c_plot_2, echo=FALSE}
sharpe2 <- function(n){
  0.85/sqrt(1 + 0.09/n)
}

tibble(sharpe.r1 = sharpe1(n), sharpe.r2 = sharpe2(n), n = n) %>%
  ggplot() + 
  geom_point(mapping = aes(x = n, y = sharpe.r1, color = "1.5")) + 
  geom_line(mapping = aes(x = n, y = sharpe.r1, color = "1.5")) +
  geom_point(mapping = aes(x = n, y = sharpe.r2, color = "2.0")) + 
  geom_line(mapping = aes(x = n, y = sharpe.r2, color = "2.0")) +
  scale_color_manual(values = c("1.5" = "blue", "2.0" = "red")) +
  labs(x = "n", y = "Sharpe Ratio", color = "Sharpe Ratio of Individual Fund",
       title = "Sharpe Ratio of the fund-of-funds") +
  theme_bw(base_size = 12)
```

The general trend in how the Sharpe Ratio changes as $n$ changes is similar in
both cases. However, all the Sharpe Ratios for each $n$ is larger for 2.0 
$\epsilon_i$-Sharpe-ratio case, when compared to 1.5 $\epsilon_i$-Sharpe-ratio 
case. To be precise, the former's Sharpe Ratio is scaled to $\frac{0.85}{0.7}$
of the latter.

\newpage

## (d)

First of all, the expected return of a single fund in this setting is:

$$
\mathbf{E}[r_i] = \beta \mathbf{E}[r_M] + \mathbf{E}[\epsilon_i]
$$

Therefore, using the properties $\beta = 0$, $\sigma_i = 0.1$ and 
$\mathbf{E}[\epsilon_i] = 1.5 \cdot \sigma_i$,

$$
\mathbf{E}[r_i - 0.01] = 0 \cdot \mathbf{E}[r_M] + 1.5 \cdot 0.1 - 0.01
= 0.14
$$

Moreover, the variance of a single fund is:

$$
\begin{aligned}
\mathbf{V}[r_i] &= \mathbf{V}[\beta r_M + \epsilon_i] \\
&= \mathbf{V}[\beta r_M] + \mathbf{V}[\epsilon_i] + 
2 Cov(\beta r_M, \epsilon_i) \\
&= \beta^2 \mathbf{V}[r_M] + \sigma_i^2 +
2 \beta Cov(r_M, \epsilon_i) \\
&= \beta^2 \mathbf{V}[r_M] + \sigma_i^2 \\
&(\because \epsilon_i \perp \!\!\! \perp r_M)
\end{aligned}
$$

Similarly, using the given properties:

$$
\mathbf{V}[r_i] = 0^2 \cdot \mathbf{V}[r_M] + 0.1^2
= 0.1^2
$$

Therefore, the Sharpe ratio of the fund becomes

$$
\frac{\mathbf{E}[r_i - 0.01]}{\sqrt{\mathbf{V}[r_i]}}
= \frac{0.14}{\sqrt{0.1^2}} = 1.4
$$

\newpage

If we plot the (d)-case and (c)-case on the same graph:

```{r q1d_plot_1, echo=FALSE}
sharpe3 <- 1.4

tibble(sharpe.r1 = sharpe1(n), n = n) %>% 
  ggplot() + 
  geom_point(mapping = aes(x = n, y = sharpe.r1, color = "Case (c)")) + 
  geom_line(mapping = aes(x = n, y = sharpe.r1, color = "Case (c)")) +
  geom_hline(yintercept = sharpe3, color = "red") +
  scale_color_manual(values = c("Case (c)" = "blue", "Case (d)" = "red")) +
  labs(x = "n", y = "Sharpe Ratio", color = "Cases",
       title = "Sharpe Ratio of the fund-of-funds",
       subtitle = "Sharpe Ratio of Inidividual Funds = 1.5") +
  theme_bw(base_size = 12)
```

The single fund seems to be much better in terms of Sharpe ratio than the 
earlier scenarios. This would mostly be attributed to the fact that the single
fund contains no market risk yet has an already high Sharpe ratio of $1.5$ by
itself.

\newpage

The comparison for when the Sharpe ratio of $\epsilon_i$ is $2.0$ would yield
a similar result with the Sharpe ratio of the fund being $1.9$

```{r q1d_plot_2, echo=FALSE}
sharpe4 <- 1.9

tibble(sharpe.r2 = sharpe2(n), n = n) %>% 
  ggplot() + 
  geom_point(mapping = aes(x = n, y = sharpe.r2, color = "Case (c)")) + 
  geom_line(mapping = aes(x = n, y = sharpe.r2, color = "Case (c)")) +
  geom_hline(yintercept = sharpe4, color = "red") +
  scale_color_manual(values = c("Case (c)" = "blue", "Case (d)" = "red")) +
  labs(x = "n", y = "Sharpe Ratio", color = "Cases",
       title = "Sharpe Ratio of the fund-of-funds",
       subtitle = "Sharpe Ratio of Inidividual Funds = 2.0") +
  theme_bw(base_size = 12)
```