---
title: "Homework 3"
subtitle: 'FINM 33210: Bayesian Statistical Inference and Machine Learning'
author: "Ki Hyun"
date: 'Due: 23:59 (CT) May 19th 2023'
output: pdf_document
---

```{r packages, include=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
```


# Problem 0.1

## (a)

$$
\begin{aligned}
\mathbf{E}[\mathbf{h}' \mathbf{r}] &=
\mathbf{E}[\sum_{i = 1}^n h_i r_i] \\
&= \sum_{i = 1}^n \mathbf{E}[h_i r_i] \\
&= \frac{1}{n} \sum_{i = 1}^n \mathbf{E}[r_i] \\
&(\because \mathbf{h} = (1/n, 1/n, \cdots, 1/n)) \\
&= \frac{1}{n} \sum_{i = 1}^n \mathbf{E}[\beta r_M + \epsilon_i] \\
&= \frac{1}{n} \sum_{i = 1}^n \left( \mathbf{E}[\beta r_M] + 
\mathbf{E}[\epsilon_i] \right) \\
&= \frac{1}{n} \sum_{i = 1}^n \mathbf{E}[\beta r_M] +
\frac{1}{n} \sum_{i = 1}^n \mathbf{E}[\epsilon_i] \\
&= \frac{\beta}{n} \sum_{i = 1}^n \mathbf{E}[r_M] +
\frac{1}{n} \sum_{i = 1}^n \mathbf{E}[\epsilon_i] \\
&= \frac{\beta}{n} n \mathbf{E}[r_M]  +
\frac{1}{n} \sum_{i = 1}^n \mathbf{E}[\epsilon_i]\\
&= \beta \mathbf{E}[r_M] +
\frac{1}{n} \sum_{i = 1}^n \mathbf{E}[\epsilon_i]
\end{aligned}
$$

$$
\begin{aligned}
\mathbf{V}[\mathbf{h}' \mathbf{r}] &=
\mathbf{V}[\sum_{i = 1}^n h_i r_i] \\
&= \mathbf{V}[\frac{1}{n} \sum_{i = 1}^n r_i] \\
&(\because \mathbf{h} = (1/n, 1/n, \cdots, 1/n)) \\
&= \frac{1}{n^2} \mathbf{V}[\sum_{i = 1}^n r_i] \\
&= \frac{1}{n^2} \mathbf{V}[\sum_{i = 1}^n (\beta r_M + \epsilon_i)] \\
&= \frac{1}{n^2} \mathbf{V}[n \beta r_M + \sum_{i = 1}^n \epsilon_i] \\
&= \frac{1}{n^2} \left( \mathbf{V}[n \beta r_M] + 
\sum_{i = 1}^n \mathbf{V}[\epsilon_i] + 
\sum_{i = 1}^n 2Cov(n \beta r_M, \epsilon_i) +
\sum_{i \neq j} 2Cov(\epsilon_i, \epsilon_j)
\right) \\
&= \frac{1}{n^2} \left( (n \beta)^2 \mathbf{V}[r_M] + 
\sum_{i = 1}^n \mathbf{V}[\epsilon_i] + 
\sum_{i = 1}^n 2n \beta Cov(r_M, \epsilon_i) +
\sum_{i \neq j} 2Cov(\epsilon_i, \epsilon_j)
\right) \\
&= \frac{1}{n^2} \left( (n \beta)^2 \mathbf{V}[r_M] + 
\sum_{i = 1}^n \mathbf{V}[\epsilon_i] + 
2 \sum_{i \neq j} Cov(\epsilon_i, \epsilon_j)
\right) \\
&(\because \epsilon_i \perp \!\!\! \perp r_M) \\
&= \frac{1}{n^2} \left( (n \beta)^2 \mathbf{V}[r_M] + 
\sum_{i = 1}^n \mathbf{V}[\epsilon_i] \right) \\
&(\because \text{ for } i \neq j, \ \epsilon_i \perp \!\!\! \perp \epsilon_j) \\
&= \beta^2 \mathbf{V}[r_M] + 
\frac{1}{n^2} \sum_{i = 1}^n \mathbf{V}[\epsilon_i] \\
&= \beta^2 \sigma^2_M + \frac{1}{n^2} \sum_{i = 1}^n \sigma^2_i
\end{aligned}
$$

Therefore, the functions $f()$ and $g()$ can be explicitly defined as:

$$
f(\beta, \sigma^2_M) := \beta^2 \sigma^2_M
$$

and

$$
g(\sigma^2_1, \cdots, \sigma^2_n) := \frac{1}{n^2} \sum_{i = 1}^n \sigma^2_i
$$

\newpage

## (b)

It was given that $\beta = 0.5$, $\sigma_M = 0.2$, and $\sigma_i \approx 0.03$.

Using the definition in (a) for $f()$ and $g()$,

$$
\frac{f}{f + g} = \frac{\beta^2 \sigma^2_M}{\beta^2 \sigma^2_M + 
\frac{1}{n^2} \sum_{i = 1}^n \sigma^2_i}
$$

Now substituting the values  $\beta = 0.5$, $\sigma_M = 0.2$, and 
$\sigma_i \approx 0.03$.

$$
\frac{f}{f + g} (n) \approx \frac{0.5^2 \cdot 0.2^2}{
0.5^2 \cdot 0.2^2 + \frac{1}{n^2} n \cdot 0.03^2}
= \frac{0.01}{0.01 + \frac{0.0009}{n}}
= \frac{1}{1 + \frac{0.09}{n}}
$$

If we numerically compute and plot $f/(f + g)$ as a function of $n$ for 
$n = 2, \dots, 30$:

```{r q1b_plot, echo=FALSE}
f_over_fg <- function(n){
  1/(1 + 0.09/n)
}
n <- seq(2, 30, by = 1)

tibble(ffg = f_over_fg(n), n = n) %>% 
  ggplot(mapping = aes(x = n, y = ffg)) + 
  geom_point() + 
  geom_line() +
  ylim(c(0.95, 1)) +
  labs(x = "n", y = "f/(f + g)",
       title = "Fraction of Variance Explained by the Market") +
  theme_bw(base_size = 12)
```

\newpage

## (c)

We know from (a) that

$$
\mathbf{E}[\mathbf{h}' \mathbf{r}] = \beta \mathbf{E}[r_M] +
\frac{1}{n} \sum_{i = 1}^n \mathbf{E}[\epsilon_i]
$$

Therefore,

$$
\begin{aligned}
\mathbf{E}[\mathbf{h}' \mathbf{r} - 0.01] &=
\mathbf{E}[\mathbf{h}' \mathbf{r}] - \mathbf{E}[0.01] \\
&= \beta \mathbf{E}[r_M] + \frac{1}{n} \sum_{i = 1}^n \mathbf{E}[\epsilon_i]
- 0.01 \\
&= \beta \cdot 0.07 + \frac{1}{n} \sum_{i = 1}^n 1.5 \cdot \sigma_i - 0.01 \\
&= 0.07 \beta + \frac{1.5}{n} \sum_{i = 1}^n \sigma_i - 0.01
\end{aligned}
$$

Moreover, we also know from (a) that

$$
\mathbf{V}[\mathbf{h}' \mathbf{r}] = \beta^2 \sigma^2_M + 
\frac{1}{n^2} \sum_{i = 1}^n \sigma^2_i
$$

Now using the assumptions of (b)

$$
\begin{aligned}
\frac{\mathbf{E}[\mathbf{h}' \mathbf{r} - 0.01]}
{\sqrt{\mathbf{V}[\mathbf{h}' \mathbf{r}]}} &=
\frac{0.07 \beta + \frac{1.5}{n} \sum_{i = 1}^n \sigma_i - 0.01}
{\sqrt{\beta^2 \sigma^2_M + \frac{1}{n^2} \sum_{i = 1}^n \sigma^2_i}} \\
& \approx \frac{0.07 \cdot 0.5 + \frac{1.5}{n} \sum_{i = 1}^n 0.03 - 0.01}
{\sqrt{0.5^2 \cdot 0.2^2 + \frac{1}{n^2} \sum_{i = 1}^n 0.03^2}} \\
&= \frac{0.07 \cdot 0.5 + 1.5 \cdot 0.03 - 0.01}
{\sqrt{0.5^2 \cdot 0.2^2 + \frac{0.03^2}{n}}} \\
&= \frac{0.35 + 0.45 - 0.1}{\sqrt{1 + \frac{0.09}{n}}} \\
&= \frac{0.7}{\sqrt{1 + \frac{0.09}{n}}}
\end{aligned}
$$

If we numerically compute and plot the Sharpe ratio as a function of $n$ for 
$n = 2, \dots, 30$:

```{r q1c_plot_1, echo=FALSE}
sharpe1 <- function(n){
  0.7/sqrt(1 + 0.09/n)
}

tibble(sharpe.r = sharpe1(n), n = n) %>% 
  ggplot(mapping = aes(x = n, y = sharpe.r)) +
  geom_point() + 
  geom_line() +
  ylim(c(0.68, 0.7))+
  labs(x = "n", y = "Sharpe Ratio",
       title = "Sharpe Ratio of the fund-of-funds across n") +
  theme_bw(base_size = 12)
```

If the Sharpe ratio of $\epsilon_i$ is $2.0$,

$$
\begin{aligned}
\frac{\mathbf{E}[\mathbf{h}' \mathbf{r} - 0.01]}
{\sqrt{\mathbf{V}[\mathbf{h}' \mathbf{r}]}} &=
\frac{0.07 \beta + \frac{2.0}{n} \sum_{i = 1}^n \sigma_i - 0.01}
{\sqrt{\beta^2 \sigma^2_M + \frac{1}{n^2} \sum_{i = 1}^n \sigma^2_i}} \\
& \approx \frac{0.07 \cdot 0.5 + \frac{2.0}{n} \sum_{i = 1}^n 0.03 - 0.01}
{\sqrt{0.5^2 \cdot 0.2^2 + \frac{1}{n^2} \sum_{i = 1}^n 0.03^2}} \\
&= \frac{0.07 \cdot 0.5 + 2.0 \cdot 0.03 - 0.01}
{\sqrt{0.5^2 \cdot 0.2^2 + \frac{0.03^2}{n}}} \\
&= \frac{0.35 + 0.6 - 0.1}{\sqrt{1 + \frac{0.09}{n}}} \\
&= \frac{0.85}{\sqrt{1 + \frac{0.09}{n}}}
\end{aligned}
$$

\newpage

Moreover, if we plot the two different $\epsilon_i$-Sharpe-ratio together:

```{r q1c_plot_2, echo=FALSE}
sharpe2 <- function(n){
  0.85/sqrt(1 + 0.09/n)
}

tibble(sharpe.r1 = sharpe1(n), sharpe.r2 = sharpe2(n), n = n) %>%
  ggplot() + 
  geom_point(mapping = aes(x = n, y = sharpe.r1, color = "1.5")) + 
  geom_line(mapping = aes(x = n, y = sharpe.r1, color = "1.5")) +
  geom_point(mapping = aes(x = n, y = sharpe.r2, color = "2.0")) + 
  geom_line(mapping = aes(x = n, y = sharpe.r2, color = "2.0")) +
  scale_color_manual(values = c("1.5" = "blue", "2.0" = "red")) +
  labs(x = "n", y = "Sharpe Ratio", color = "Sharpe Ratio of Individual Fund",
       title = "Sharpe Ratio of the fund-of-funds") +
  theme_bw(base_size = 12)
```

The general trend in how the Sharpe Ratio changes as $n$ changes is similar in
both cases. However, all the Sharpe Ratios for each $n$ is larger for 2.0 
$\epsilon_i$-Sharpe-ratio case, when compared to 1.5 $\epsilon_i$-Sharpe-ratio 
case. To be precise, the former's Sharpe Ratio is scaled to $\frac{0.85}{0.7}$
of the latter.

\newpage

## (d)

Using the same assumptions in (c), for the 1.5 $\epsilon_i$-Sharpe-ratio case,
the new Sharpe Ratio of the fund of funds becomes:

$$
\begin{aligned}
\frac{\mathbf{E}[\mathbf{h}' \mathbf{r} - 0.01]}
{\sqrt{\mathbf{V}[\mathbf{h}' \mathbf{r}]}} &=
\frac{0.07 \beta + \frac{1.5}{n} \sum_{i = 1}^n \sigma_i - 0.01}
{\sqrt{\beta^2 \sigma^2_M + \frac{1}{n^2} \sum_{i = 1}^n \sigma^2_i}} \\
&= \frac{0.07 \cdot 0 + \frac{1.5}{n} \sum_{i = 1}^n 0.1 - 0.01}
{\sqrt{0^2 \cdot 0.2^2 + \frac{1}{n^2} \sum_{i = 1}^n 0.1^2}} \\
&= \frac{1.5 \cdot 0.1 - 0.01}{\sqrt{\frac{0.1^2}{n}}} \\
&= \frac{0.14}{0.1} \sqrt{n} \\
&= 1.4 \sqrt{n}
\end{aligned}
$$

If we plot the (d)-case and (c)-case on the same graph:

```{r q1d_plot_1, echo=FALSE}
sharpe3 <- function(n){
  1.4*sqrt(n)
}

tibble(sharpe.r1 = sharpe1(n), sharpe.r2 = sharpe3(n), n = n) %>% 
  ggplot() + 
  geom_point(mapping = aes(x = n, y = sharpe.r1, color = "Case (c)")) + 
  geom_line(mapping = aes(x = n, y = sharpe.r1, color = "Case (c)")) +
  geom_point(mapping = aes(x = n, y = sharpe.r2, color = "Case (d)")) + 
  geom_line(mapping = aes(x = n, y = sharpe.r2, color = "Case (d)")) +
  scale_color_manual(values = c("Case (c)" = "blue", "Case (d)" = "red")) +
  labs(x = "n", y = "Sharpe Ratio", color = "Cases",
       title = "Sharpe Ratio of the fund-of-funds",
       subtitle = "Sharpe Ratio of Inidividual Funds = 1.5") +
  theme_bw(base_size = 12)
```

Simply investing in a single fund case of (d) seems to be better in terms
of Sharpe ratio when individual fund's Sharpe Ratio is $1.5$. The difference
in Sharpe ratio between the two cases is greater for larger $n$ values.

\newpage

Now when individual fund's Sharpe Ratio is $2.0$:

$$
\begin{aligned}
\frac{\mathbf{E}[\mathbf{h}' \mathbf{r} - 0.01]}
{\sqrt{\mathbf{V}[\mathbf{h}' \mathbf{r}]}} &=
\frac{0.07 \beta + \frac{2.0}{n} \sum_{i = 1}^n \sigma_i - 0.01}
{\sqrt{\beta^2 \sigma^2_M + \frac{1}{n^2} \sum_{i = 1}^n \sigma^2_i}} \\
&= \frac{0.07 \cdot 0 + \frac{2.0}{n} \sum_{i = 1}^n 0.1 - 0.01}
{\sqrt{0^2 \cdot 0.2^2 + \frac{1}{n^2} \sum_{i = 1}^n 0.1^2}} \\
&= \frac{2.0 \cdot 0.1 - 0.01}{\sqrt{\frac{0.1^2}{n}}} \\
&= \frac{0.19}{0.1} \sqrt{n} \\
&= 1.9 \sqrt{n}
\end{aligned}
$$

Moreover,

```{r q1d_plot_2, echo=FALSE}
sharpe4 <- function(n){
  1.9*sqrt(n)
}

tibble(sharpe.r1 = sharpe1(n), sharpe.r2 = sharpe4(n), n = n) %>% 
  ggplot() + 
  geom_point(mapping = aes(x = n, y = sharpe.r1, color = "Case (c)")) + 
  geom_line(mapping = aes(x = n, y = sharpe.r1, color = "Case (c)")) +
  geom_point(mapping = aes(x = n, y = sharpe.r2, color = "Case (d)")) + 
  geom_line(mapping = aes(x = n, y = sharpe.r2, color = "Case (d)")) +
  scale_color_manual(values = c("Case (c)" = "blue", "Case (d)" = "red")) +
  labs(x = "n", y = "Sharpe Ratio", color = "Cases",
       title = "Sharpe Ratio of the fund-of-funds",
       subtitle = "Sharpe Ratio of Inidividual Funds = 2.0") +
  theme_bw(base_size = 12)
```

Similarly, simply investing in a single fund case of (d) seems to be better in 
terms of Sharpe ratio when individual fund's Sharpe Ratio is $2.0$ as well. The 
difference in Sharpe ratio between the two cases is greater when individual 
funds have $2.0$ Sharpe ratio than when individual funds have $1.5$ Sharpe 
ratio. Within the above graph, the difference in Sharpe ratio between the two
cases is greater for larger $n$ values.